---
title: "Regional Training on Big Data and Data Science for Gender Statistics in Asia and the Pacific"
author: "Your Name (Your Institution)"
subtitle: '*An automated report on gender statistics*'
date: "`r Sys.Date()`"
output:
  html_document:
    toc: yes
    toc_float: true
    code_folding: hide
  pdf_document:
    toc: yes
  word_document:
    toc: no
---


```{r libraries, include = FALSE}
# Load the required libraries

# For manipulating data
library(tidyverse)

# For tables display
library(flextable)

```


```{r setup, include=FALSE}
# Set echo=FALSE as default for code cells
# The code folding for HTLM output is thus deactivated
knitr::opts_chunk$set(echo = TRUE)

```


# Data analysis
We provide here some basic elements about the data set and descriptive statistics on some variables of interest.  

* First we load the data
* second we start exploring some variables

```{r echo =FALSE}
# GenderData <- read.csv("https://www.unsiap.or.jp/on_line/2024/RAP/Data/GenderData.csv")
GenderData <- read.csv("Data/SG_GEN_PARL_ap.csv")
```


we use the SDG 5.5.1 indicator (*SG_GEN_PARL_ap.csv*) available as a CSV file from 
[UNSSTAT database](https://unstats.un.org/SDGS/Indicators/Database/).  We loaded the data set from UNSIAP server and compile the report on `r Sys.Date()`.

## Data exploration

The data set contains   **`r  nrow(GenderData)`  observations**, and  **`r  ncol(GenderData)`** variables.

Here is a list of all variables available in the data set: 

```{r}
names(GenderData)
```

Let's see also the list of countries. This information is in the variable *GeoAreaName*. 

```{r}
ListCountries <- unique(GenderData$GeoAreaName)
ListCountries

```
We have `r length(ListCountries)` countries in our data set, over 11 years (2021-2020)

## Panel data: Countries over time

The variables starting with *X* report the "*Proportion of seats held by women in national parliaments*" in percentage.  For example, in 2020, we can compute some summary statistics of that variable, over all countries.  


```{r  }
# Report for each country in 2020

StatsByCountry <- GenderData %>%
  group_by(GeoAreaName) %>%
  summarise(Percentage = X2020) %>%
  arrange(desc(Percentage))

```



```{r}
# Table display
# To reduce the homogenization efforts and the number of functions to be called, 
# it is recommended to define formatting properties once and for all

# set_flextable_defaults(
#   font.size = 10,   table.layout = "autofit")

# Printing the table
flextable(StatsByCountry)
```

# As a graphic: 

Since we have a good number of countries, a bar chart would probably describe the data in a very good way. We follow here some good practices in terms of data visualization: 

- The bar chart is **horizontal** to facilitate the reading of the name of countries
- The bar chart is **ordered** to follow the data (largest on top)

```{r }
# Plot the percentages by country
StatsByCountry  %>%
  ggplot()+
  aes(x = reorder(GeoAreaName, Percentage), y = (Percentage)) +
  geom_bar(stat = "identity", 
             fill= "lightsteelblue1") +
  labs(title ="Proportion of seats held by women in national parliaments ", 
       subtitle = "Year 2020", 
       caption  = paste("Source: SDG, (file date: ",Sys.Date(), ")"), 
       x= "Countries",
       y = "Percentages") +
  coord_flip() +
  theme_minimal()

```

```{r}
knitr::knit_exit()
```


# Some extensions on this dynamic report

##  Selection of the Year

The previous analysis was done for a specific year (2020). Imagine now the same report but for any year in [2010, 2020]. let's pick a year in the code chunck below! 

```{r}
MyYear <- 2019
```

> we have now selected the year **`r MyYear`**, let's see the ***new*** report!

### As a table

We use the same table as before. Maybe that could be a function later ;-).

```{r  }
# We first need to identify the variable that we will summarize:
SelectedColumn <- paste0("X", MyYear)

# Now only for a specific year
StatsByCountryYear <- GenderData %>%
  group_by(GeoAreaName) %>%
  summarise(Percentage = .data[[SelectedColumn]]) %>%
  arrange(desc(Percentage))

```


```{r}
flextable(StatsByCountryYear, title=" sdlkfsf")
```


As expected, the graphic shows the difference with the 

```{r }
# Plot the exports for the specific year
ExportsByType_Year  %>%
  ggplot(aes(x = reorder(Principle_Exports, total), y = (total))) +  
    geom_bar(stat = "identity", color="white", fill= "lightsteelblue1") +
      labs( x= "Exports categories", 
            y = "Value (Vatu)", 
            title = paste("Exports from Vanuatu"),
            subtitle = paste(" Selected year:", MyYear), 
            caption  = paste("Source: NSO, (file date: ",Sys.Date(), ")"), ) +
  coord_flip() +
  theme_minimal()

```





