---
title: 'An (automated) Report on Gender Statistics'
subtitle: "*Regional Training on Big Data and Data Science for Gender Statistics in Asia and the Pacific*"
author: "Christophe Bontemps (Your Institution)"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: yes
    toc_float: true
    code_folding: hide
  pdf_document:
    toc: yes
  word_document:
    toc: no
---


```{r libraries, include = FALSE}
# Load the required libraries, 
# if these libraries are not already installed on your computer:
#  - Use the package installation from the bottom right panel in RStudio (recommended)
# - Or (experimented users) uncomment the next line (only once)
#install.packages(c("tidyverse", "flextable"))

# For manipulating data
library(tidyverse)

# For tables display
#library(flextable)

```


```{r setup, include=FALSE}

# Set echo=FALSE as default for code cells
# The code folding for HTLM output is thus deactivated
knitr::opts_chunk$set(echo = TRUE)

```

We provide here some basic elements about the data set and descriptive statistics on some variables of interest.  

* First we load the data (directly from GitHub)
* Second we start exploring some variables

```{r echo =FALSE}
GenderData <- read.csv("https://raw.githubusercontent.com/XtopheB/BigDataForGenderStats/refs/heads/main/Data/SG_GEN_PARL_ap.csv")

# If you want, you can store also the data locally and load it  from there
# GenderData <- read.csv("Data/SG_GEN_PARL_ap.csv")

```


We will work with data on the SDG 5.5.1 indicator ("*Proportion of seats held by women in national parliaments*") available as a CSV file from  [UNSSTAT database](https://unstats.un.org/SDGS/Indicators/Database/). For easing the exercise, we stored this data set on GitHub server and use a direct download of the data set from there (*SG_GEN_PARL_ap.csv*). This report was compiled on **`r format(Sys.Date(), "%A %d %B")`**.


# Overall data exploration

> The data set contains   **`r  nrow(GenderData)`  observations**, and  **`r  ncol(GenderData)`** variables.

Here is a list of all variables available in the data set: 

```{r}
names(GenderData)
```

Let's see also the list of countries. This information is in the variable *GeoAreaName*. 

```{r}
ListCountries <- unique(GenderData$GeoAreaName)
ListCountries

```
We have `r length(ListCountries)` countries in our data set, over 11 years (2021-2020)

# Seats held by women in parliaments 

The variables starting with *X* report the "*Proportion of seats held by women in national parliaments*" in percentage.  For example, **in 2020**, we can compute some summary statistics of that variable, over all countries, using the *X2020* variable.   


```{r  }
# Report for each country in 2020

StatsByCountry <- GenderData %>%
  group_by(GeoAreaName) %>%
  summarise(Percentage = X2020) %>%  # here, we selected X2020 (Year 2020)
  arrange(desc(Percentage))

```

## As a table
We can represent the indicator as a sorted table, with highest values being on top.  

```{r}
# Table display (using flextable package)

# Printing the table
flextable(StatsByCountry)
```
It seems that some values are missing (bottom of the table) and have been replaced by zeros. This issue should be addressed in the analysis, but we will not do it in this report.  

## As a graphic 

Since we have a good number of countries, a bar chart would probably describe the data in a very good way. We follow here some good practices in terms of data visualization: 

- The bar chart is **horizontal** to facilitate the reading of the name of countries
- The bar chart is **ordered** to follow the data (largest on top)

```{r }
# Plot the percentages by country
StatsByCountry  %>%
  ggplot()+
  aes(x = reorder(GeoAreaName, Percentage), y = (Percentage)) +
  geom_bar(stat = "identity", 
             fill= "lightsteelblue1") +
  labs(title ="Proportion of seats held by women in national parliaments ", 
       subtitle = "Year 2020", 
       caption  = paste("Source: UNSSTAT (file date: ",Sys.Date(), ")"), 
       x= "Countries",
       y = "Percentages") +
  coord_flip() +
  theme_minimal()

```

# Conclusion
 In this report, we explored the Proportion of seats held by women in national parliaments for **`r nrow(GenderData)`** countries, as a table and as a graphic for the specific year selected, here **2020**.

To change the year, we should change all text and code concerning 2020 and effortlessly make another report for a different year.  We could also - and this will be done in another report - have all code and text written as function of a global variable, so that changing that value at the very beginning of the code would consistently change the whole report. 








